---
title: "The dependencies conundrum"
post_date: 2018-06-06
layout: single
permalink: /dependencies-conundrum/
categories: r-blog-en
output: jekyllthat::jekylldown
excerpt_separator: <--!more--> 
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.path="../assets/img/")
```

A little dive into CRAN package dependencies.

## What is a dependencies? 

Roughly, a dependencie is a package needed by another package to work: in other words, if someone has already coded a function that you need, you can use it by listing another package as a dependency. But that also implies that whenever someone install your package, the dependencies are installed too. 

## Where dependencies can be problematic 

### Heavy dependencies

Some R packages are heavy, and can be complex to install on specific machine. For example, `{readr}` does not install easily on small-RAM ubuntu machine: [see this issue](https://github.com/tidyverse/readr/issues/544)] for more information about that. That leads to not being able to install this package, but also other packages dependending on `{readr}`: a behavior I noticed lately with the `{rtweet}` package when I tried to install it on a Digital Ocean small ubuntu droplet: [see this issue](https://github.com/mkearney/rtweet/issues/246).

### The clipr effect

Whenever one of your dependencies is removed from the CRAN, your package will be removed also. That happens to a series of package recently, notably the `{reprex}` 📦, but that's food for another blogpost [see this issue](https://github.com/mdlincoln/clipr/issues/30) if you want to read more in the meantime. 

In a few words, just imagine that if A depends on B, B on C and C on D, if ever the D package is removed from the CRAN, you'll need to find a solution to keep your package up to date. 

## Dependencies when developping package

### How many dependencies? 

It's a tough question and everybody has its take on it. Mine is: try to avoid them if:

+ The dependency has a lot of other dependencies / is heavy to install 
+ You just need one easy to rewrite function from that package

To answer the first question, you can use : 

```{r depends, message = FALSE}
library(miniCRAN)
pkgDep("attempt", suggests = FALSE)
pkgDep("proustr", suggests = FALSE)
```
Which means that at least all these packages will be installed whenever I want to install one of these. 

## Diving into the CRAN 

Let's answer the second question by having a look at how many package from the CRAN have just one function imported from another package. 

There would be several way to do that: but I chosed to parse the NAMESPACE files, as it is the place where `Depends` and `Imports` are listed with the same tag. Maybe not the most easy to deploy solution, but I'll be happy to hear about other solution! 

```{r eval = FALSE}
library(rvest)
library(purrr)
library(glue)

# Get the NAMESPACE file

extract_namespace <- function(package, exdir){
  cat(crayon::green(glue("Extracting {package} NAMESPACE to {exdir} at {Sys.time()}.")), sep = "\n")
  a <- read_html(glue("https://cran.r-project.org/web/packages/{package}/index.html")) %>%
    html_nodes("a") %>%
    html_attr("href") %>%
    keep(~ grepl("tar.gz", .x)) %>%
    gsub("\\.\\./\\.\\./\\.\\.", "https://cran.r-project.org", .)
  
  tmp <- tempfile(fileext = ".tar.gz")
  download.file(a, tmp)
  nspc <- grep("NAMESPACE", untar(tmp, list = TRUE), value = TRUE)
  untar(tmp, file = nspc, exdir = exdir)
  unlink(tmp)
}

# Parsing of the import fields
parse_import <- function(pkg, vec){
  cleaned <- gsub('\\(|,|\\)|"', " ", vec) 
  splitted_bis <- strsplit(cleaned, " ") %>% as_vector() %>% discard(~ nchar(.x) == 0)
  type <- splitted_bis[1]
  dep <- splitted_bis[2]
  fun <- splitted_bis[3:length(splitted_bis)]
  if ( grepl("From", cleaned) ) {
    data.frame(origin = pkg, 
               type = type, 
               dep = dep, 
               fun = fun, 
               stringsAsFactors = FALSE)
  } else {
    data.frame(origin = pkg,
               type = type, 
               dep = dep, 
               fun = NA, 
               stringsAsFactors = FALSE)
  }
}

# Parsing the full file
parse_namespace <- function(pkg, file){
  readLines(normalizePath(file)) %>%
    keep(~ grepl("import", .x)) %>%
    map2_df(pkg, ~ parse_import(.y, .x))
}

# Wrap all this into one
extract_pkg_and_df_namespace <- function(list, exdir = "nsp"){
  walk(list, extract_namespace, exdir = exdir)
  df <- data.frame(pkg = list.files(exdir), 
                   loc = list.files(exdir, recursive = TRUE, full.names = TRUE), 
                   stringsAsFactors = FALSE)
  map2_df(df$pkg, df$loc, parse_namespace)
}

# Get the package db
pkg_db <- tools::CRAN_package_db()

# FIRE 🛫 
res <- extract_pkg_and_df_namespace(pkg_db$Package)
```

So, don't try this at home, but know it took around 4 hours for me to retrieve all these data. 
### Let's explore 

```{r include = FALSE}
res <- readRDS("~/Seafile/documents_colin/R/allnamespace/res.RDS")
```

Here, we 

```{r}
library(tidyverse)
fun_topo <- res %>%
  filter(type == "importFrom") %>%
  group_by(origin, dep) %>%
  summarise(exported_fun = n()) %>%
  ungroup()

fun_topo %>%
  arrange(desc(exported_fun))

fun_topo %>%
  filter(exported_fun >= 50) %>%
  arrange(desc(exported_fun)) 

fun_topo %>%
  filter(exported_fun < 30) %>%
  ungroup() %>%
  count(exported_fun) %>%
  ggplot() + 
  aes(exported_fun, n) + 
  geom_col()

just_one <- fun_topo %>%
  filter(exported_fun == 1) %>%
  left_join(res)

nrow(just_one)

just_one %>% 
  count(dep, fun, sort = TRUE)

just_one %>%
  count(origin)

just_one %>%
  filter(origin == "abbyyR")

res %>%
  filter(fun == "read_csv")
```

