---
title: "Solving #AdventOfCode day 3 and 4 with R"
post_date: 2018-12-02
layout: single
permalink: /advent-code-3-4/
categories: r-blog-en
output: jekyllthat::jekylldown
excerpt_separator: <!--more-->
---

```{r setup, include = FALSE}
library(knitr)
opts_chunk$set(fig.path="../assets/img/", eval = TRUE, error =TRUE)
```
 
Solving the puzzles of [Advent of Code](https://adventofcode.com) with R.

 <!--more-->
 
[Disclaimer] Obviously, this post contains a big spoiler about Advent of Code, as it gives solutions for solving day 1 and 2. 

## Advent of Code 

> Advent of Code is an Advent calendar of small programming puzzles for a variety of skill sets and skill levels that can be solved in any programming language you like. [About Advent of Code](https://adventofcode.com/2018/about)

## Day three, part one

The first part of the challenge for day 3 starts well: we have to read in R this file: 

```{r}
readLines("input3.txt", n = 6)
```

It's a description of "spots", claimed by the elves, inside a big square piece of fabric (supposed to be for making Santa's clothes (remember we're coding with Christmas theme ;) ). 

It reads:

```
"#1 @ 483,830: 24x18" 
Id / distance from left / distance from top / width / height 
```

In other words, the "claim" with ID one is for a spot that starts at 483 inches from the left of the fabric and at 830 inches from the top. This piece is 24 inches wide and 18 inches tall. 

> How many square inches of fabric are within two or more claims?

Our challenge is to determine how many squares of fabric are claimed more than once. In over words, if we take all the squares from all the elves, how many square inches are claimed by more than one elf? 

Let's start by reading the file: 

```{r message=FALSE}
library(tidyverse)
df <- read.delim("input3.txt", header = FALSE) %>%
  as.tibble() %>%
  print() %>%
  # Clean the column 
  mutate(V1 = str_replace_all(V1, "[#@,:x]", " "),
         V1 = str_replace_all(V1, " {2,}", " "), 
         V1 = str_replace_all(V1, "^ ", "")) %>%
  # Separate the column
  separate(V1, into = c("id", "left", "top", "wide", "tall")) %>% 
  # Make sure everything is numeric
  mutate_all(as.numeric) %>%
  # Create the right and down element
  mutate(
    right = left + wide, 
    down = top + tall 
  ) 
cat("Cleaned version:\n")
df
```

So now that we have that, how can we get **all the squares claimed** more than once? To do this, we'll go row-wise on this data.frame. With each row, we'll start by create vectors which correspond to all the claimed squares. To do that, we'll create two vectors: row and columns, use expand.grid to create all the possible combination (i.e all the locations on the fabric) for this claim.

For example, imagine a claim that goes from square 1 to 2 and from line 3 to 4, we'll create a vector 1:2 and 3:4, and when combining that we'll have the id for all claimed squares: 1x3, 1x4, 2x3 and 2x4. We'll do this for all the claims, then count the number of occurrence for each newly created id.

```{r}
res <- map_df(
  1:nrow(df), 
  function(x){
    res <- list(
     col = (df$left[x] + 1):(df$left[x]+df$wide[x]),
     row = (df$top[x] + 1):(df$top[x]+df$tall[x])
    )
    expand.grid(res$col, res$row) %>%
      mutate(claim_id = df$id[x]) %>%
      unite(col = "idx", c("Var1", "Var2"), sep = "x")
  }
) %>% as.tibble()
res
res %>% 
  count(idx) %>%
  filter(n > 1) %>%
  count()  
```

## Day three, part two

> What is the ID of the only claim that doesn't overlap?

The second part is a little bit different: we need to identify the claim which doesn't overlap with any other claim. In other words, the claim that contains only square inches which are claimed once. 

So how to do that ? 

We'll start from our previous table and count the number of square by claim: 

```{r}
n_of_claims_by_id <- res %>% 
  count(claim_id)
n_of_claims_by_id
```

Now we'll count the number of time each square is claimed:

```{r}
n_of_claims_by_inch <- res %>%
  count(idx)
n_of_claims_by_inch
```


Let's add this info to our result object: 

```{r}
res <- res %>%
  left_join(n_of_claims_by_inch, by = "idx") 
res
```

Here, we can read that square at `484x831` has been claimed three times (`n`), and notably by the ID `#1`. So let's filter on the square that has only been claimed once:

```{r}
res <- res %>%
  filter(n == 1)
res
```

Now for the final trick: a claim with no overlap would have, in our result, the same number of occurrence in the `claim_id` column from `res` and in the `idx` from `n_of_claims_by_inch`

```{r}
res %>%
  count(claim_id) %>%
  left_join(n_of_claims_by_id, by = "claim_id") %>%
  filter(nn == n)
```

So here it is: the claim_id `1124` is the one we are looking for! 

## Day four, part one

This one was also fun to do: we have another weird data file, that we  need to parse. This data is said to record the date and time, all days between 00 and 00:59 at a manufacturing lab, when the guard is awake of asleep. 

```{r}
library(lubridate)
day4 <- read.delim("input4.txt", header = FALSE) %>% as.tibble()
day4

time_minute <- day4 %>%
  # Create a hour / event column
  separate(V1, into = c("hour", "event"), sep = "] ") %>% 
  mutate(hour = gsub("\\[", "", hour), 
         hour = ymd_hm(hour)) %>%
  # Order the hour column
  arrange(hour) %>%
  # Create a column with the guard ID id present
  mutate(guard_id = str_extract(event, "#([0-9]*)"), 
         guard_id = str_replace(guard_id, "#", ""), 
         no_id = is.na(guard_id), 
         record = 1:n()
         ) %>%
  # Add the guard ID everywhere
  fill(guard_id) %>%
  # Filter to keep only the event rows
  filter(no_id) %>% 
  select(-no_id) %>%
  # Spread for event and hour
  spread(event, hour) %>%
  # Fill the wakes up column so that there is no NA
  fill(`wakes up`, .direction = "up") %>%
  select(-record) %>%
  # Remove the line with NA in the fall asleep column
  drop_na(`falls asleep`) %>%
  # Get one minutes less, as the wakes up minute is considered to be an
  # awaken minute
  mutate(`wakes up` = `wakes up` - minutes(1),
  # Create a list column with the sequence of minutes the guard is asleep
         minutes_asleep = map2(`falls asleep`, `wakes up`, 
                               seq, by = "min"),
  # Compute how many time the guard is asleep
         sleepy_time = `wakes up` - `falls asleep`) 
time_minute
```

> Find the guard that has the most minutes asleep. What minute does that guard spend asleep the most?

So, our first task is to know which guard is asleep the most and on which minute of the hour. Let's first compute who this guard is:

```{r}
which_guard <- time_minute %>%
  group_by(guard_id) %>%
  summarise(s = sum(sleepy_time)) %>%
  top_n(1, s) 
```

Now we know who he is, let's see at which minute of the hour he's the most asleep:

```{r}
time_minute %>%
  filter(guard_id == which_guard$guard_id) %>%
  unnest(minutes_asleep) %>%
  print() %>%
  # Extract the minute part
  mutate(minutes_asleep = minute(minutes_asleep)) %>%
  count(minutes_asleep, sort = TRUE)%>%
  top_n(1)
```

## Day four, part one

The second part was quite straightforward once you've got the previous answer. 
> Of all guards, which guard is most frequently asleep on the same minute?

```{r}
time_minute %>%
  unnest(minutes_asleep) %>%
  mutate(minutes_asleep = minute(minutes_asleep)) %>%
  count(guard_id, minutes_asleep, sort = TRUE) %>% 
  top_n(1)
```

