---
title: "Why do we use arrow as an assignment operator?"
post_date: 2018-07-30
layout: single
permalink: /r-assignment/
categories: r-blog-en
output: jekyllthat::jekylldown
excerpt_separator: <!--more-->
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.path="../assets/img/", eval = TRUE, error =TRUE)
```
 
A [Twitter Thread](https://twitter.com/_ColinFay/status/1006139974377443328) turned into a blog post.
 
 <!--more-->
 
In June, I published a little [thread](https://twitter.com/_ColinFay/status/1006139974377443328) on Twitter about the history of the `<-` assignment operator in R. Here is a blog post version of this thread. 

## Historical reasons

As you all know, R comes from S. But you might not know a lot about S (I don't). This language used `<-` as an assignment operator. It's partly because it was inspired by a language called APL, which also had this sign for assignment. 

But why again? APL was designed on a specific keyboard, which had a key for `<-`:

<img src ="https://upload.wikimedia.org/wikipedia/commons/thumb/9/9f/APL-keybd2.svg/410px-APL-keybd2.svg.png">

At that time, it was also chosen because there was no `==` for testing equality: equality was tested with `=`, so assigning a variable needed to be done with another symbol. 

<img align = "center" src = "/assets/img/blog/apl.jpg">

> From [APL Reference Manual](http://www.softwarepreservation.org/projects/apl/Books/APL360ReferenceManual)

Until [2001](http://developer.r-project.org/equalAssign.html), in R, `=` could only be used for assigning function arguments, like `fun(foo = "bar")` (remember that R was born in 1993). So before 2001, the `<-` was the standard (and only way) to assign value into a variable.  

Before that, `_` was also a valid assignment operator. It was removed in [R 1.8](https://cran.r-project.org/src/base/NEWS.1):

<img align = "center" src = "/assets/img/blog/runderscore.jpg">

(So no, at that time, no snake_case_naming_convention)

Colin Gillespie published [some of his code from early 2000](https://www.jumpingrivers.com/blog/r-from-the-turn-of-the-century/), where assignment was made like this :) 

The main reason "equal assignment" was introduced is because other languages uses `=` as an assignment method, and because it increased compatibility with S-Plus. 

## And today?

### Readability

Nowadays, there are seldom any cases when you can’t use one in place of the other. It’s safe to use `=` almost everywhere. Yet, `<-` is preferred and advised in R Coding style guides: 

+ https://google.github.io/styleguide/Rguide.xml#assignment
+ http://adv-r.had.co.nz/Style.html

One reason, if not historical, to prefer the `<-` is that it clearly states in which side you are making the assignment (you can assign from left to right or from right to left in R): 

```{r}
a <-  12
13 -> b 
a
b
a -> b
a <- b
```

The RHS assignment can for example be used for assigning the [result of a pipe](https://rud.is/b/2015/02/04/a-step-to-the-right-in-r-assignments/)

```{r message = FALSE}
library(dplyr)
iris %>%
  filter(Species == "setosa") %>% 
  select(-Species) %>%
  summarise_all(mean) -> res
res
```

Also, it's easier to distinguish equality comparison and assignment in the last line of code here:

```{r}
c <- 12
d <- 13
e = c == d
f <- c == d
```

Note that `<<-` and `->>` also exist: 

```{r}
create_plop_pouet <- function(a, b){
  plop <<- a
  b ->> pouet
}
create_plop_pouet(4, 5)
plop
pouet
```

And that Ross Ihaka uses `=` : <https://www.stat.auckland.ac.nz/~ihaka/downloads/JSM-2010.pdf>

### Environments

There are some environment and precedence differences. For example, assignment with `=` is only done on a functional level, whereas `<-` does it on the top level when called inside as a function argument.

```{r error = TRUE}
median(x = 1:10)
x
median(x <- 1:10)
x
```

In the first code, you're passing `x` as the parameter of the `median` function, whereas the second one is creating a variable x in the environment, and uses it as the first argument of `median`. Note that it works because `x` is the name of the parameter of the function, and won't work with `y`:

```{r}
median(y = 12)
median(y <- 12)
```

There is also a difference in parsing when it comes to both these operators (but I guess this never happens in the real world), one failing and not the other:

```{r error = TRUE}
x <- y = 15
x = y <- 15
c(x, y)
```

It is also good practice because it clearly indicates the difference between function arguments and assignation:

```{r}
x <- shapiro.test(x = iris$Sepal.Length)
x
```

And this weird behavior: 

```{r}
rm(list = ls())
data.frame(
  a = rnorm(10),
  b <- rnorm(10)
)
a
b
```

## Little bit unrelated but 

I love this one: 

```{r}
g <- 12 -> h
g
h
```

Which of course is not doable with `=`. 

## Other operators

Some users pointed out on Twitter that this could make the code a little bit harder to read if you come from another language. `<-` is use "only" use in F#, OCaml, R and S (as far as Wikipedia can tell). Even if `<-` is rare in programming, I guess its meaning is quite easy to grasp, though. 

Note that the second most used assignment operator is `:=` (`=` being the most common). It's used in `{data.table}` and `{rlang}` notably. The `:=` operator is not defined in the current R language, but has not been removed, and is still understood by the R parser. You can't use it on the top level:

```{r}
a := 12
```

But as it is still understood by the parser, you can use `:=` as an infix without any %%, for assignment, or for anything else: 

```{r}
`:=` <- function(x, y){
  x$y <- NULL
  x
}
head(iris := Sepal.Length)
```

You can see that `:=` was used as an assignment operator <https://developer.r-project.org/equalAssign.html> :

> All the previously allowed assignment operators (<-, :=, _, and <<-) remain fully in effect 

Or in R NEWS 1: 

<img align = "center" src = "/assets/img/blog/colonequal.png">

## See also 

+ Around 29': https://channel9.msdn.com/Events/useR-international-R-User-conference/useR2016/Forty-years-of-S
+ [Use = or <- for assignment?](http://blog.revolutionanalytics.com/2008/12/use-equals-or-arrow-for-assignment.html)
+ [What are the differences between “=” and “<-” in R?](https://stackoverflow.com/questions/1741820/what-are-the-differences-between-and-in-r/1742550#1742550)
+ [Assignment Operators](https://stat.ethz.ch/R-manual/R-devel/library/base/html/assignOps.html)
