---
title: "A quick #WorldEmojiDay exploration"
post_date: 2018-07-17
layout: single
permalink: /worldemojiday
categories: r-blog-en
output: jekyllthat::jekylldown
excerpt_separator: <--!more--> 
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.path="../assets/img/", eval = TRUE)
```

Let's celebrate #WorldEmojiDay with a quick exploration of my [own twitter account](https://twitter.com/_ColinFay). 

## The 📦

We'll need: 

#### From Github

+ `{emo}`

```{r eval = FALSE}
remote::install_github("hadley/emo")
```

#### From CRAN

+ `{dplyr}`
+ `{tidyr}`
+ `{rtweet}`
+ `{tidytext}`


> Note: This page has been created at:

```{r}
Sys.time()
```

## The 🔍

Let's get my last 3200 tweets:

```{r}
library(emo)
library(rtweet)
library(dplyr)
res <- get_timeline(
  "_ColinFay",
  n = 3200
)
names(res)
```

Here is what the `text` column looks like: 

```{r}
res %>% 
  pull(text) %>%
  .[1:5]
```

As you can see, the emojis are not printed in the console, but converted to weird characters like `\U0001f4e6` and such. These are unicode characters: translations of the emojis into a language your machine can understand. I won't go deeper into this, here are two resources you can read if you want to know more about encoding: 

+ [The Absolute Minimum Every Software Developer Absolutely, Positively Must Know About Unicode and Character Sets (No Excuses!)](https://www.joelonsoftware.com/2003/10/08/the-absolute-minimum-every-software-developer-absolutely-positively-must-know-about-unicode-and-character-sets-no-excuses/)

+ [What emojis tell us about encoding](https://www.codenewbie.org/blogs/what-emojis-tell-us-about-encoding)

## The 📊

Let's use the `{emo}` package to extract the emojis from the text. Inspired by `{stringr}`, this package has a `ji_extract_all` function that is designed to __extract all the emojis from a character vector__. We'll use it on out text column, then extract the date and emo column. We then pass the result to `tidyr::unnest` in order to remove the empty emo rows (i.e, the tweets without an emoji). 

```{r}
library(tidyr)
emos <- res %>%
  mutate(
    emo = ji_extract_all(text)
  ) %>%
  select(created_at,emo) %>%
  unnest(emo)

emos

emos %>%
  count(emo, sort = TRUE)
```

So apparently, I use a lot of 🤔. But also talk about 📦, which sounds more appropriate :)

As you can see, `{tibble}` converts elements to emojis when printing. When using a data.frame, you have a simple unicode translation: 

```{r}
emos %>%
  as.data.frame() %>%
  head()
```

## The 🏷

Let's flag all the emojis with their names: 

```{r}
emos %>%
  left_join(
    data.frame(
      emo = ji_name, 
      name = names(ji_name)
    )
  ) %>% 
  count(emo, name, sort = TRUE)
```

## The 🔠

And finally, let's see what are the most associated words with the emojis we just saw:

```{r}
library(tidytext)
emos_with_id <- res %>%
  mutate(
    emo = ji_extract_all(text)
  ) %>% 
  select(status_id, text, emo) %>%
  tidyr::unnest(emo)

emos_with_id %>%
  unnest_tokens(word,text) %>%
  anti_join(stop_words) %>%
  anti_join(proustr::stop_words) %>%
  anti_join(
    data.frame(
      word = c("https", "t.co", "https", "gt")
    )
  ) %>% 
  count(emo, word, sort = TRUE)
```

And what are the most used emojis with "rstats"? 

```{r}
emos_with_id %>%
  unnest_tokens(word, text) %>%
  anti_join(stop_words) %>%
  anti_join(proustr::stop_words) %>%
  anti_join(
    data.frame(
      word = c("https", "t.co", "https", "gt")
    )
  ) %>%
  count(emo, word, sort = TRUE) %>%
  filter(
    word == "rstats"
  )
```

## Other cool functions

I recently discovered the `ji_glue()` function which allows you to insert an emoji easily into a character vector : 

```{r}
ji_glue("I love to code :package:")
ji_glue("Sometimes they make me :scream:")
ji_glue("Sometimes they make me :cry:")
ji_glue("Sometimes they make me :fear:")
ji_glue("But in the end I'm always :tada:")
```

The `ji()` function can also be used inside your markdown, so you can write: 

"I hate *backtick* r emo::ji("bug") *backtick*", and it will come as: "I hate `r emo::ji("bug")`".

(of course, replace *backtick* by actuwith backticks :) ).

## That's all folks 🎬

That's all for today! Now have a nice emoji day `r emo::ji("tada")`
