---
title: "An Introduction to Docker for R Users"
post_date: 2019-01-06
layout: single
permalink: /docker-r-reproducibility/
categories: r-blog-en
output: jekyllthat::jekylldown
excerpt_separator: <!--more-->
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.path="../assets/img/", eval = TRUE, error =TRUE)
```
 
A quick introduction on using Docker for reproducibility in R. 

<!--more-->

> Disclaimer: this blog post is an introduction to Docker for beginners, and will takes some shortcuts ;) 

## What is Docker? 

Docker is "a computer program that performs operating-system-level virtualization, also known as 'containerization'" [Wikipedia](https://en.wikipedia.org/wiki/Docker_(software)). As any first line of a Wikipedia article about tech, this sentence is obscure to anyone not already familiar with the content of the article. 

So, to put it more simply, __Docker is a program that allows to manipulate (launch and stop) multiple operating systems (called containers) on your machine (your machine will be called the host)__. Just imagine having 10 RaspberryPi with different flavors of Linux, each focused on doing one simple thing, that you can turn on and off whenever you need to ; but all of this happens on your computer. 

## Why Docker & R? 

Docker is designed to enclose environments inside an image / a container. What this allows, for example, is to have a Linux machine on a Macbook, or a machine with R 3.3 when your main computer has R 3.5. Also, this means that __you can use older versions of a package for a specific task, while still keeping the package on your machine up-to-date__. 

This way, you can "solve" dependencies issues: if ever you are afraid dependencies will break your analysis when packages are updated, build a container that will **always** have the software versions you desire: be it Linux, R, or any package. 

## Docker images vs Docker containers

On your machine, you're going to need two things: images, and containers. __Images are the definition of the OS, while the containers are the actual running instances of the images__. You'll need to install the image just once, while the containers are to be launched whenever you need this instance. And of course, multiple containers of the same images can be run at the same time. 
 
To compare with R, this is the same principle as installing vs loading a package: a package is to be downloaded once, while it has to be launched every time you need it. And a package can be launched in several R sessions at the same time easily.

## Dockerfile

A Docker image is built from a `Dockerfile`. This file is the configuration file, and describes several things: from what previous docker image you are building this one, how to configure the OS, and what happens when you `run` the container. In a sense, it's a little bit like the `DESCRIPTION` + `NAMESPACE` files of an R package, which describes which are the dependencies to your package, gives meta information, and states which functions and data are to be available to the users `library()`ing the package. 

So, let's build a **very basic** `Dockerfile` for R, focused on reproducibility. The idea is this one: I have today an analysis that works (for example contained in a `.R` file), and I want to be sure this analysis will always work in the future, regardless of any update to the packages used. 

So first, create a folder for your analysis, and a Dockerfile:

```
mkdir ~/mydocker
cd ~/mydocker
touch Dockerfile
```

### `FROM`

Every `Dockerfile` starts with a `FROM`, which describes what image we are building our image from. There are a lot of official images, and you can also build from a local one.

This `FROM` is, in a way, describing the dependency of your image ; just as in R, when building a package, you always rely on another package (be it only the `{base}` package). 

If you're going for an R based image, Dirk Eddelbuettel & Carl Boettiger
 are maintaining [rocker](https://hub.docker.com/u/rocker), a collection of Docker images for R you can use. We'll use the `rocker/r-base` in this blogpost.
 
```
FROM rocker/r-base
```

### `RUN`

Once we've got that, we'll add some `RUN` statements: these are commands which mimic command line commands. Remember what we want: an image that will, ad vitam aeternam, run an analysis as if we were still today. So what we'll do is use the `{checkpoint}` package. 

The command to make R execute something, from the terminal, is `R -e "my code"`. Let's add a `{checkpoint}` installation. 

```
FROM rocker/r-base

RUN R -e "install.packages('checkpoint')"
```

We need a `/root/.checkpoint` folder to use `{checkpoint}`, let's create that one with `mkdir` (make directory).

```
FROM rocker/r-base

RUN R -e "install.packages('checkpoint')"

RUN mkdir /root/.checkpoint
```

### `COPY`

Now, I need to get the script for my analysis from my machine (host) to the container. For that, we'll need to use `COPY localfile pathinthecontainer`. I'll first create a folder to receive everything, with `mkdir`. Note that here, the `myscript.R` has to be in the same folder as the `Dockerfile` on your computer.

Let's say this is the content of `myscript.R`: 

```{r eval = FALSE}
library(checkpoint)
checkpoint("2019-01-06")
library(tidystringdist)
df <- tidy_comb_all(iris, Species)
p <- tidy_stringdist(df)
write.csv(p, "p.csv")
```

Here, the `{tidystringdist}` that will be installed in the machine will be the one from the date of today, even if I build this image in one year, or two, or four.

```
FROM rocker/r-base

RUN R -e "install.packages('checkpoint')"

RUN mkdir /home/analysis

COPY myscript.R /home/analysis/myscript.R
```

### `CMD`

`CMD` is the command to be run every time you'll launch the docker. What we want is `myscript.R` to be sourced. 

```
FROM rocker/r-base

RUN R -e "install.packages('checkpoint')"

RUN mkdir /home/analysis

COPY myscript.R /home/analysis/myscript.R

CMD R -e "source('/home/analysis/myscript.R')"
```

## Build, and run

### Build

Now, go and build your image. From your terminal, in the directory where the Dockerfile is located, run: 

```
docker build -t analysis .
```

`-t name` is the name of the image (here analysis), and `.` means it will build the `Dockerfile` in the current working directory. 

### `run` 

Then, just launch with:

```
docker run analysis 
```

And your analysis will be run 🎉!

## Export container content

One thing to do now: you want to access what is created by your analysis (here `p.csv`) outside your container ; i.e, on the host. Because yes, as for now, everything that happens in the container stays in the container. __So what we need is to make the docker container share a folder with the host. For this, we'll use what is called Volume, which are (roughly speaking), a way to tell the Docker container to use a folder from the host as a folder inside the container__.

That way, everything that will be created in the folder by the container will persist after the container is turned off. To do this, we'll use the -v flag when running the container, with `path/from/host:/path/in/container`. Also, create a folder to receive the results in both : 

```
FROM rocker/r-base

RUN R -e "install.packages('checkpoint')"

RUN mkdir /home/analysis && mkdir /home/results

COPY myscript.R /home/analysis/myscript.R

CMD cd /home/analysis && R -e "source('myscript.R')" && mv /home/analysis/p.csv /home/results/p.csv
```

```
mkdir ~/mydocker/results 
docker run -v ~/mydocker/results:/home/results  analysis 
```

Wait for the computation to be done, and... 

```
ls ~/mydocker/results  
p.csv
```
🤘

## What to do next? 

So now, every time you'll launch this Docker image, the analysis will be performed and you'll get the result back. With no problem of dependencies: the packages will always be installed from the day you desire. Although, this can be a little bit long to run as the packages are installed each time you run the container. But as I said in the Disclaimer, this is a basic introduction to Docker, R and reproducibility, so the goal was more to get beginners on board with Docker :) 

Other things you can do would be: 

+ Using [`{packrat}`](https://rstudio.github.io/packrat/), and get the library bundle in the container. 

+ Use `remotes::install_version()` if you want your analysis to be based on package version instead of a time based installation. 


```
FROM rocker/r-base

RUN R -e "install.packages('remotes'); remotes::install_version('tidystringdist', '0.1.2')"

...
```

+ Use the Volume trick to bring data **into** your container, so that any data will be analysed in the very same environment. 

And other cool stuffs, but that's for another blog post ;) 

## Know more about Docker

Some resources:

+ [rocker official website](https://www.rocker-project.org/)
+ [An Introduction to Rocker: Docker Containers for R](https://journal.r-project.org/archive/2017/RJ-2017-065/index.html)
+ [Noam Ross - Docker for the UseR](https://github.com/noamross/nyhackr-docker-talk)
+ [Production-ready R: Getting started with R and docker](https://www.youtube.com/watch?v=lfG8cTqRRNA)
+ [Applications with R and Docker](https://www.youtube.com/watch?v=JOIKy_89c-o)
+ [Docker for beginners](https://docker-curriculum.com/)