---
title: "JavaScript cont in R"
post_date: 2019-09-23
layout: single
permalink: /js-const-r/
categories: r-blog-en
output: jekyllthat::jekylldown
excerpt_separator: <!--more-->
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.path="../assets/img/", eval = TRUE, error =TRUE)
```

One thing I like about JavaScript is the `const` declaration method, which allows you to declare a variable one time, and that variable can't be reassigned after that. I.e, this piece of code will throw an error:

```{bash}
node -e "const x = 12; x  = 14"
```

The cool thing about this is that you can't override the variable by mistake: once it's set, it's set. On the other hand, R allows you to override almost any variable (well, except some reserved variables).

I asked Twitter if there was any implementation of that concept in R. The use case, for example, would arise when you have a value that takes some time to compute. If I do my computation, I can accidentally override it later on. Event more if you're using notebook, where you create symbols and values all along your document.

```{r eval = FALSE}
a <- some_very_complex_computation()
# [...] Going on the weekend
a <- "Hello there!"
```

Here, I have no way to prevent myself from erasing the value in `a`. Of course, there are always rigor, explicit variable name, and not-assigning-things-without-thinking but you know how it is in the real world, and there is no Cmd + Z there. 

[Romain](https://twitter.com/romain_francois/status/1176092011084046336) pointed out that `?lockBinding` existed, and that it was what I was looking for. And that does.

Here's how it works: it takes a character string referring to a symbol, and an environment, and prevents from assigning any new value to this symbol in the given environment. 

```{r error = TRUE}
x <- 12
lockBinding("x", .GlobalEnv)
x <- 13
```

And here's a small wrapper to do that:

```{r error = TRUE}
lock <- function(x){
  lockBinding(
    deparse(
      substitute(x)), 
    env = parent.frame()
  )
}

plop <- 12
lock(plop)
plop <- 13

pouet <- function(){
  plop <- 14
  print(plop)
  lock(plop)
  plop <- 13
}
pouet()
```

So there I could do 

```{r eval = FALSE}
a <- some_very_complex_computation()
lock(a)
# [...] Going on the weekend
a <- "Hello there!"
```

And there, I have prevented myself from erasing my `a` variable. 
Of course, it's not the same as JavaScript `const`, as there is always a way to unlock the symbol. 

```{r error = TRUE}
x <- 12
lock(x)
x <- 13
```

```{r}
unlockBinding("x", .GlobalEnv)
x <- 13
x
```

But I think it's a rather elegant solution for preventing yourself from unwanted variable overwriting. 

See also:

+ [Make a Constant in R with Active Binding](https://iqis.netlify.com/post/2019/07/22/how-to-make-a-constant-in-r/)

Some answers to the Twitter thread also suggested using R6... but that will be for another post :) 
