---
title: "2018 through {cranlogs}"
post_date: 2019-01-02
layout: single
permalink: /12-months-cranlogs/
categories: r-blog-en
output: jekyllthat::jekylldown
excerpt_separator: <!--more-->
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.path="../assets/img/", eval = TRUE, error =TRUE)
```
 
2018 at glance with `{cranlogs}`.

<!--more-->

Let's load the necessary packages. 

```{r}
library(cranlogs)
library(data.table)
library(lubridate)
library(ggplot2)
library(magrittr)
```

## All downloads

We'll use `{cranlogs}` to retrieve the data from the RStudio CRAN mirror.  

First, the number of package downloads by day in 2018.

```{r}
total_dl <- cran_downloads(from = "2018-01-01", to = "2018-12-31")

# Turn to a data.table
setDT(total_dl)

# Round the date to month and week
total_dl[, `:=`(
  round_week = floor_date(date, "week" ),
  round_month = floor_date(date, "month" )
  ) ]
```

How many download in total?

```{r}
total_dl[, .(total = sum(count))]
```

Let's plot this:

```{r}
random_viridis <- function(n){
  sample(viridis::viridis(100), n)
}
total_dl[, .(count = sum(count)), round_week] %>%
  ggplot(aes(round_week, count)) + 
  geom_col(fill = random_viridis(1)) + 
  labs(
    title = "Packages downloads by Week on RStudio CRAN mirror", 
    subtitle = "data via {cranlogs}", 
    x = "week"
  ) + 
  theme_minimal()

total_dl[, .(count = sum(count)), round_month] %>%
  ggplot(aes(round_month, count)) + 
  geom_col(fill = random_viridis(1)) + 
  labs(
    title = "Packages downloads by Month on RStudio CRAN mirror", 
    subtitle = "data via {cranlogs}", 
    x = "month"
  ) + 
  theme_minimal()
```

## R download

Let's now have a look at the number of downloads for R itself: 

```{r}
total_r <- cran_downloads("R", from = "2018-01-01", to = "2018-12-31")

setDT(total_r)

total_r[, `:=`(
  round_week = floor_date(date, "week" ),
  round_month = floor_date(date, "month" )
) ]

```

How many download in total?

```{r}
total_r[, .(total = sum(count))]
```

Plotting this: 

```{r}
total_r[, .(count = sum(count)), round_week] %>%
  ggplot(aes(round_week, count)) + 
  geom_col(fill = random_viridis(1)) + 
  labs(
    title = "R downloads by Week on RStudio CRAN mirror", 
    subtitle = "data via {cranlogs}", 
    x = "week"
  ) + 
  theme_minimal()

total_r[, .(count = sum(count)), round_month] %>%
  ggplot(aes(round_month, count)) + 
  geom_col(fill = random_viridis(1)) + 
  labs(
    title = "R downloads by Month on RStudio CRAN mirror", 
    subtitle = "data via {cranlogs}", 
    x = "month"
  ) + 
  theme_minimal()
```

Let's have a look to the number of download by R version:

```{r}
total_r[, .(count = sum(count)), version][order(count, decreasing = TRUE)] %>%
  head(10)

total_r[, .(count = sum(count)), version][order(count)] %>%
  head(10)

total_r[, .(count = sum(count)), version][order(count, decreasing = TRUE)] %>%
  head(10) %>% 
  ggplot(aes(reorder(version, count), count)) +
  coord_flip() +
  geom_col(fill = random_viridis(1)) + 
  labs(
    title = "10 most downloaded R versions in 2018 on RStudio CRAN mirror", 
    subtitle = "data via {cranlogs}", 
    x = "version"
  ) + 
  theme_minimal()
```

And by os:

```{r}
total_r[, .(total = sum(count)), os]
```

And a happy new year 🎉🎉
