---
title: "Advent of Code 2019-01 with R & JavaScript"
post_date: 2019-12-01
layout: single
permalink: /aoc-2019-01/
categories: r-blog-en
output: jekyllthat::jekylldown
excerpt_separator: <!--more-->
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.path="../assets/img/", eval = TRUE, error =TRUE)
bubble::set_node_engine()
```

Solving Advent of Code 2019-01 with R and JavaScript. 

[Disclaimer] Obviously, this post contains a big spoiler about Advent of Code, as it gives solutions for solving day 1.

[Disclaimer bis] I'm no JavaScript expert so this might not be the perfect solution. TBH, that's also the case for the R solution. 

## About the JavaScript code

The JavaScript code has been written in the same RMarkdown as the R code. It runs thanks to the `{bubble}` package: https://github.com/ColinFay/bubble

## Instructions

Find the instructions at: <https://adventofcode.com/2019/day/1>

## R solution

### Part one

```{r}
# Read
ipt <- read.delim( "input1.txt",  header = FALSE )
# Get the sum of each element, divided by 3, rounded down, and substracted 2
sum( floor( ipt$V1 / 3) - 2 )
```

### Part two 

> Using a recursive function: https://en.wikipedia.org/wiki/Recursion_(computer_science)

```{r}
floorish <- function(x, start = 0){
  loc <- floor( x / 3) - 2
  if (loc > 0){
    start <- start + loc
    floorish(loc, start)
  } else {
    return(start)
  } 
}
sum( purrr::map_dbl(ipt$V1, floorish) )
```

## JS solution

### Part one & Two

```{node results = 'hide'}
var fs = require('fs')
```

> // Reading the file

```{node results = "hide"}
var res = fs.readFileSync("input1.txt", 'utf8').split("\n").filter(x => x.length != 0);
```

> // Turning to integer

```{node results = "hide"}
res = res.map(x => parseInt(x));
```

> // Doing the floor of division less 2

```{node results = "hide"}
var val = res.map(x => Math.floor(x / 3) - 2);
```

> // Suming

```{node results = "hide"}
var add = (x, y) => x + y;
```

> // Solution

```{node}
val.reduce(add)
```

> // Creating the recursive function

```{node results = "hide"}

function floorish(val, start = 0){
  loc = Math.floor(val / 3) - 2;
  if (loc > 0){
    start += loc;
    return floorish(loc, start);
  } else {
    return start;
  } 
}

```

> // Doing the computation

```{node}
res.map( x => floorish(x) ).reduce( add ) 
```
