---
title: "How to use your Garmin watch to tell your team you're going for a run"
post_date: 2020-01-08
layout: single
permalink: /api-call-garmin-watch/
categories: r-blog-en
output: jekyllthat::jekylldown
excerpt_separator: <!--more-->
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.path="../assets/img/", eval = TRUE, error =TRUE)
```

Building an API in NodeJS and R to send message to Slack from your Garmin watch. 

## Why on earth

ThinkR is a remote company, meaning that we all work from our home. 
On top of other cool things about remote work, this allows me to skip my lunch break and take a one hour break in the middle of the afternoon for sport. 
And I usually go for a run around 2 or 3 pm, but that moment in the day is not exactly the same every day. 
And most of the time, I forget to tell everyone that I'm leaving the office. 
I recently joked that my 2020 resolution was that I'll be more strict about telling when I arrive and leave the "office". 

I was sure it can be done straight from my watch. 
And guess what, it can! 

## The Slack Part

The slack API is pretty amazing and allows you to use a personal web hook and a curl call to send messages to a selected channel on Slack. 

Note: there are several packages in R that can be used to send messages to Slack, for example `{slackr}`: https://github.com/hrbrmstr/slackr & `{slackteams}`: https://github.com/yonicd/slackteams.
But as I just wanted to make a simple, unique call, it was more straightforward to write it directly. 

So: 

+ Go to `https://api.slack.com/`

+ Click on _Start Building_

+ Add an app name and add it to a workspace

<div align = "center">

<img src = "/assets/img/slackapi1.png">

</div>

+ Add a new "Incoming Webhooks", and select the Channel to post in

<div align = "center">

<img src = "/assets/img/slackapi2.png">

</div>

And Tadaa 🎉 you now have a curl call that looks like this:

```
curl -X POST -H 'Content-type: application/json' 
  --data '{"text":"Hello, World!"}' 
  https://hooks.slack.com/services/YOUR/WEBHOOK/URL/HERE
```

Now time to turn this into an API.

## Node API

Here is a very simple API built in NodeJS: 

```{javascript, eval = FALSE}
const express = require('express')
const app = express()

app.get('/', function (req, res) {
  const request = require('request');
  
  const options = {
    url: 'https://hooks.slack.com/services/YOUR/WEBHOOK/URL/HERE',
    json: true,
    body: {
      text: "I'm off for a run!"
    }
  };
  
  request.post(options);
  
  res.send('OK')
})

app.listen(9999, function () {
  console.log('API listening on port 9999!')
})
```

## R API

And with R: 

```{r eval = FALSE}
library(plumber)

#* @get /
function() {
  httr::POST(
    url = 'https://hooks.slack.com/services/YOUR/WEBHOOK/URL/HERE', 
    body = list(
      text  = "I'm off for a run!"
    ), 
    encode = "json"
  )
  
}
```

## Adding this to the watch 

I discovered that Garmin has a widget called "API calls", that let you enter an API endpoint, and the API call is done from the watch. 

https://apps.garmin.com/en-US/apps/ac9a81ab-a52d-41b3-8c14-940a9de37544

<div align = "center">

<blockquote class="twitter-tweet"><p lang="en" dir="ltr">I just discovered that I can send API calls from my Garmin watch and I&#39;m very excited about this and that also made me realized that I&#39;m definitely a big nerd. <a href="https://t.co/2h4fk8tCnf">pic.twitter.com/2h4fk8tCnf</a></p>&mdash; Colin Fay 🤘 (@_ColinFay) <a href="https://twitter.com/_ColinFay/status/1214556262660071427?ref_src=twsrc%5Etfw">January 7, 2020</a></blockquote> <script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>

</div>

So here it is, I've got a Widget on my watch that I can use to send message on Slack 🎉 

<div align = "center">

<blockquote class="twitter-tweet"><p lang="en" dir="ltr">You know you&#39;re a nerd when you deploy an API that posts to Slack from your Garmin watch so that you can tell your team that you&#39;re going for a run instead of just... you know... typing it. <a href="https://t.co/f2cwn1pxd7">pic.twitter.com/f2cwn1pxd7</a></p>&mdash; Colin Fay 🤘 (@_ColinFay) <a href="https://twitter.com/_ColinFay/status/1214655430951092224?ref_src=twsrc%5Etfw">January 7, 2020</a></blockquote> <script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>

</div>

