---
title: "Automate the Creation of an API Wrapper package by Scraping its Online Documentation"
post_date: 2020-03-12
layout: single
permalink: /fun-from-api-doc/
categories: r-blog-en
output: jekyllthat::jekylldown
excerpt_separator: <!--more-->
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.path="../assets/img/", eval = TRUE, error =TRUE, comment = "")
```

I've recently been building a package to interact with the [Rocket.chat](https://rocket.chat/) API. 
One thing I don't like about building API wrapper is that it's very repetitive: you have to copy and paste things over and over again, for each endpoint of the API. 
Arguments, documentation... 

But here's a fun idea, how about scraping the help page of the API to generate the whole thing?

## General idea

So the global skeleton I need is the one for a `.R` file for a function, with `{roxygen2}` skeleton, and as much of the function as possible. 
Then, I need the arguments to the function, which are the available parameters of the API call. 

If they are required by the API, I'll leave them without a default value. If they are not, I'll pass them `NULL` as a default parameter. 
Then, I'll create a list with these params, remove the `NULL` with an internal fun, and turn them to `JSON`. 
Then I'll be building the `{httr}` call. 

One thing I know upfront is that I won't be able to generate the whole thing, but let's try to do as much as I can!

## The code

I'll try to build a function page for a call to `"https://rocket.chat/docs/developer-guides/rest-api/users/create/"`

```{r}
library(rvest)
library(magrittr, warn.conflicts = FALSE)
library(purrr, warn.conflicts = FALSE)
page <- read_html("https://rocket.chat/docs/developer-guides/rest-api/users/create/") 
```

When on the page, we will find the args and their doc on the second table

```{r}
fun_args <- page %>% 
  html_table() %>% 
  purrr::pluck(2)
fun_args
```

Nice! We can start by creating some roxy tags out of this. 

```{r}
# Small util functions
tag_app <- function(var) {
  sprintf("#' %s", var)
}

ap_em_rox <- function(val){
  val %<>% c(
    "#' "
  )
  val
}
```

The title of the function is at the second h2 of the page:

```{r}
page %>% 
  html_nodes("h1") %>%
  pluck(2) %>%
  html_text()
```

Let's add it.

```{r}
title <- page %>% 
  html_nodes("h1") %>%
  pluck(2) %>%
  html_text()
fun_code <- title %>%
  tag_app() %>% 
  ap_em_rox()
cat(fun_code, sep = "\n")
```

And use the first paragraph as a function description.

```{r}
fun_code %<>% c(
  page %>% 
    html_nodes("p") %>% 
    pluck(1)%>%
    html_text() %>%
    gsub("\n", ". ", .) %>%
    tag_app() %>% 
    ap_em_rox()
)

cat(fun_code, sep = "\n")

```

Now, something our API doc doesn't have, the connection token we'll be passing from the auth.

```{r}
fun_code %<>% c("#' @param token The token to connect to the app.") 
styler::style_text(fun_code)
```

Ok remember our table of parameters? Let's us it now to create the roxygen tags.

```{r}
fun_code %<>% c(
  purrr::pmap_chr(
    fun_args,
    ~ {
      sprintf(
        "#' @param %s %s %s", 
        thinkr::clean_vec(..1), ..4, ..3
      )
    }
  )
) %<>% ap_em_rox()
cat(fun_code, sep = "\n")
```

Of course we'll export the function: 

```{r}
fun_code %<>% c(
  tag_app("@export")
)
```

Some deps

```{r}
fun_code %<>% c(
  tag_app("@importFrom httr POST GET add_headers content stop_for_status"),
  tag_app("@importFrom jsonlite toJSON")
  
)
```

Now, let's create the function and its arguments.

```{r}
fun_code %<>% c(
  sprintf(
    "%s <- function(tok,",
    thinkr::clean_vec(title)
  )
)

fun_code %<>% c(
  purrr::pmap_chr(
    fun_args,
    ~ {
      if (
        grepl("Optional", ..3)
      ){
        sprintf(
          "  %s = NULL,", 
          thinkr::clean_vec(..1)
        )
      } else {
        sprintf(
          "  %s,", 
          thinkr::clean_vec(..1)
        )
      }
      
    }
  )
)

fun_code[
  length(fun_code)
  ] %<>% gsub(",", "", .)
cat(fun_code, sep = "\n")
```

Ok now a little bit of the internals: 

```{r}
fun_code %<>% c(
  "){", 
  " ", 
  "  params <- list("
)

fun_code %<>% c(
  purrr::pmap_chr(
    fun_args,
    ~ {
      sprintf(
        "    %s = %s,", 
        thinkr::clean_vec(..1), thinkr::clean_vec(..1)
      )
    }
  )
)

fun_code[
  length(fun_code)
  ] %<>% gsub(",", "", .)

fun_code %<>% c(
  ")", 
  "", 
  "params <- no_null(params)", 
  "", 
  "params <- toJSON(params, auto_unbox = TRUE)"
)
cat(fun_code, sep = "\n")
```

A piece of the `{httr}` request. 
Note that our token contains a `$url` that contains the url of the chat, and the `authToken` and `userId`.

The endpoint is on the first table, as is the http method.

```{r}
method <- page %>%
  html_table() %>% 
  pluck(1)
method["HTTP Method"] 
method["URL"] 
```


```{r}

if ( method["HTTP Method" ] == "POST"){
  fun_code %<>% c(
    sprintf(
      "res <- httr::%s(", 
      method["HTTP Method"]
    ), 
    "add_headers('Content-type' = 'application/json',",
    "'X-Auth-Token' = tok$data$authToken,",
    "'X-User-Id' = tok$data$userId",
    "),",
    sprintf("url = paste0(tok$url, '%s'),", method["URL"] ),
    "body = params"
  )
} else {
  fun_code %<>% c(
    sprintf(
      "res <- httr::%s(", 
      method["HTTP Method" ]
    ), 
    "add_headers('Content-type' = 'application/json',",
    "'X-Auth-Token' = tok$data$authToken,",
    "'X-User-Id' = tok$data$userId",
    "),",
    sprintf("url = paste0(tok$url, '%s')", method["URL"] )
  )
}
```

End this:

```{r}
fun_code %<>% c(
  ")", 
  " ",
  "stop_for_status(res)", 
  "content(res)",
  "}"
) 
```

Let's see what this looks like:

```{r}
styler::style_text(fun_code)
```

And to automate even more:

```{r eval = FALSE}
usethis::use_r(title)
write(fun_code, sprintf("R/%s.R", title))
```

Pretty cool! 
Seems like I have everything... or almost. 
Let's automate it even more:

```{r}
user_methods <- read_html("https://rocket.chat/docs/developer-guides/rest-api/users/")
user_methods %>% 
  html_nodes("a") %>%
  html_attr("href") %>% 
  grep("rest-api/users/.*/", ., value = TRUE)
```

`r emo::ji("tada")`, now we have all the endpoints! Wrapping the previous code in a function, and we can automate even more. 

Wanna see the created package? 
Here is the beginning: [ColinFay/rrocketchat](https://github.com/ColinFay/rrocketchat), the function that creates all other file can be found at [data-raw/DATASET.R](https://github.com/ColinFay/rrocketchat/blob/master/data-raw/DATASET.R)
