---
title: "clientsdb - A docker image with clients comments"
post_date: 2020-07-13
layout: single
permalink: /clients-db/
categories: r-blog-en
output: jekyllthat::jekylldown
excerpt_separator: <!--more-->
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.path="../assets/img/", eval = TRUE, error =TRUE, comment = "")
```

Have you ever been looking for a ready to use database while doing training? 
Search no more, this docker is an image with a client review database dropped inside a postgre, to be used for teaching.

## About the dataset

The titles and comments are extracted from this [Google Drive](https://drive.google.com/drive/folders/0Bz8a_Dbh9Qhbfll6bVpmNUtUcFdjYmF2SEpmZUZUcVNiMUw1TWN6RDV3a0JHT3kxLVhVR2M) link that contains "amazon_review_full_csv.tar.gz", which I discovered on the [Amazon review database](https://www.kaggle.com/bittlingmayer/amazonreviews) Kaggle page.
Then the two columns date and name being randomly generated in R. 

Here is the coded used to generate the full table:

```r
library(data.table)
dataset <- fread("data/train.csv", header = FALSE, sep = ",")
names(dataset) <- c("score", "title", "comment")

nms <- paste(
  sample(charlatan:::person_en_us$first_names, nrow(dataset), TRUE), 
  sample(charlatan:::person_en_us$last_names, nrow(dataset), TRUE)
)

date <- sample(0:as.numeric(as.POSIXct("2010-01-01")), nrow(dataset), TRUE)
date <- as.POSIXct(date, origin = "1970-01-01")

dataset[
  , `:=`(
    score = NULL,
    name = nms, 
    date = date
  )
]   

data.table::fwrite(dataset, "datasetwithusers.csv")
```

## Launch and use

The main purpose of this image is to provide a "real life" tool for teaching databases use.

Info: 

- the `POSTGRES_DB` used is `clients`
- the `POSTGRES_PASSWORD` is `verysecretwow`
- the `POSTGRES_USER` is `superduperuser`

To launch the db, do: 

```{bash, eval = FALSE}
# Might take some time to warm up
docker run --rm -d -p 5432:5432 --name clientsdb colinfay/clientsdb:latest
``` 

Then, for example from R: 

```{r}
library(DBI)

con <- dbConnect(
  RPostgres::Postgres(),
  dbname = 'clients', 
  host = 'localhost',
  port = 5432, 
  user = 'superduperuser',
  password = 'verysecretwow'
)

dbListTables(con)
res <- dbSendQuery(con, "SELECT score, title, name, date FROM clients LIMIT 5")
dbFetch(res)
dbClearResult(res)

res <- dbSendQuery(con, "SELECT title, name, date FROM clients WHERE date = '1998-05-12' LIMIT 10")
dbFetch(res)
dbClearResult(res)

dbDisconnect(con)
```

And then stop the db.

```{bash, eval = FALSE}
docker stop clientsdb 
``` 

