---
title: "Advent of Code 2020-01 with R & JavaScript"
post_date: 2020-12-01
layout: single
permalink: /aoc-2020-01/
categories: r-blog-en
output: jekyllthat::jekylldown
excerpt_separator: <!--more-->
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.path="../assets/img/", eval = TRUE, error =TRUE, warning = FALSE)
bubble::set_node_engine()
```

Solving Advent of Code 2020-01 with R and JavaScript. 

[Disclaimer] Obviously, this post contains a big spoiler about Advent of Code, as it gives solutions for solving day 1.

[Disclaimer bis] I'm no JavaScript expert so this might not be the perfect solution. TBH, that's also the case for the R solution. 

## About the JavaScript code

The JavaScript code has been written in the same RMarkdown as the R code. It runs thanks to the `{bubble}` package: https://github.com/ColinFay/bubble

## Instructions

+ Combine all the possible duos and trios of the input numbers

+ Locate the combination that equals 2020

+ Multiply the numbers from this combination to get the puzzle answer

Find the complete instructions at: <https://adventofcode.com/2020/day/1>. 

## R solution

### Part one

```{r}
# Read
ipt <- read.delim( "2020-01-aoc.txt",  header = FALSE )
library(dplyr, warn.conflicts = FALSE)
# Creating all the combinations
tidystringdist::tidy_comb_all(ipt$V1) %>%
  # Adding them together
  mutate(sum = V1 + V2) %>%
  # keeping only the one that equals 2020
  filter(sum == 2020) %>%
  # Multiplying it
  mutate(res = V1 * V2) %>%
  # Getting the answer
  pull(res)
```

### Part two 

```{r}
# Creating all the combinations
combn(
  ipt$V1, 
  3
) %>% 
  # Transposing the matrix and turning it to a tibble
  t() %>%
  as_tibble() %>%
  # Same old song as step 1
  mutate(sum = V1 + V2 + V3) %>%
  filter(sum == 2020)  %>%
  mutate(res = V1 * V2 * V3) %>%
  pull(res)
```

## JS solution

### Part one & Two

> // Reading the file and converting it to int

```{node results = 'hide'}
const fs = require('fs')
var ipt = fs.readFileSync("2020-01-aoc.txt", 'utf8').split("\n").filter(x => x.length != 0);
ipt = ipt.map(x => parseInt(x));
```

> // Doing the combn

```{node results = "hide"}
var ipt2 = ipt;
var comb = ipt.flatMap(i => ipt2.map(j => [i, j]))
```

> // Getting the 2020

```{node results = "hide"}
var twentytwenty = comb.filter(x => (x[0] + x[1]) === 2020)
```

> // Solution

```{node}
twentytwenty.reduce(y => y[0] * y[1])
```

> // Same with three

```{node results = "hide"}
var ipt3 = ipt;
var comb2 = ipt.flatMap(i => ipt2.flatMap(j => ipt3.map(k => [i, j, k])));
```

> // Doing the computation

```{node results = "hide"}
var twentytwenty2 = comb2.filter(x => (x[0] + x[1] + x[2]) === 2020)
```

> // Solution

```{node}
twentytwenty2.map(y => y[0] * y[1] * y[2])[0]
```