---
title: "Advent of Code 2020-04 with R"
post_date: 2020-12-04
layout: single
permalink: /aoc-2020-04/
categories: r-blog-en
output: jekyllthat::jekylldown
excerpt_separator: <!--more-->
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.path="../assets/img/", eval = TRUE, error =TRUE, warning = FALSE)
bubble::set_node_engine()
```

Solving Advent of Code 2020-04 with R and JavaScript. 

[Disclaimer] Obviously, this post contains a big spoiler about Advent of Code.

## Instructions

### Step 1

+ Inputs are of the form of

```
ecl:gry pid:860033327 eyr:2020 hcl:#fffffd
byr:1937 iyr:2017 cid:147 hgt:183cm

iyr:2013 ecl:amb cid:350 eyr:2023 pid:028048884
hcl:#cfa07d byr:1929

hcl:#ae17e1 iyr:2013
eyr:2024
ecl:brn pid:760753108 byr:1931
hgt:179cm

hcl:#cfa07d eyr:2025 pid:166559648
iyr:2011 ecl:brn hgt:59in
```

Where each "passport" is separated by a new line.

To be considered valid, a passport has to have all of `c("byr","iyr","eyr", "hgt", "hcl", "ecl", "pid")`.

### Step 2

To be considered valid, a passport has to have all of `c("byr","iyr","eyr", "hgt", "hcl", "ecl", "pid")`, plus, each entry has to validate a series of rules.

Find the complete instructions at: <https://adventofcode.com/2020/day/4>. 

## R solution

### Part one

```{r}
# Read the data
ipt <- readLines("2020-04-aoc.txt" ) %>% 
  # pasting everything into one big character string
  paste(collapse = "\n") %>% 
  # splitting where there are two \n\n (new lines)
  strsplit("\n\n") %>% 
  .[[1]] %>%
  # Removing the new lines
  gsub("\n", " ", .)

library(purrr)

is_north_pole_valid <- function(
  x, 
  patt = c("byr","iyr","eyr", "hgt", "hcl", "ecl", "pid")
){
  map_lgl(
    patt, 
    ~ grepl(.x, x)
  ) %>% all()
}

map_dbl(
  ipt, ~{
    if (is_north_pole_valid(.x)) return(1)
    return(0)
  }
) %>% sum()
```

### Part two 

```{r}
ipt %>%
  # Remove the unvalid passport
  discard(~ !is_north_pole_valid(.x)) %>%
  # Split stuff
  str_split(pattern = " ") %>%
  map_dbl(~{
    # Now, the heavy lifting
    # we have a string of key:value pairs, 
    # split it
    vals <- map_chr(.x, ~ gsub("([^:]*):(.*)", "\\2", .x)) 
    names(vals) <- map_chr(.x, ~ gsub("([^:]*):(.*)", "\\1", .x)) 
    
    # checking that values are inside range
    if (! dplyr::between(vals["byr"], 1920, 2002) ) return(0)
    if (! dplyr::between(vals["iyr"], 2010, 2020) ) return(0)
    if (! dplyr::between(vals["eyr"], 2020, 2030) ) return(0)
    # If height is in inch, check the range
    if ( grepl("in", vals["hgt"]) ) {
      if (! dplyr::between(gsub("in", "", vals["hgt"]), 59, 76)) return(0)
    } else if (grepl("cm", vals["hgt"])) {
      if (! dplyr::between(gsub("cm", "", vals["hgt"]), 150, 193)) return(0)
    } else {
      # No unit provided
      return(0)
    }
    if (! grepl("^#[a-f0-9]{6}$", vals["hcl"])) return(0)
    if (! vals["ecl"] %in% c("amb","blu", "brn", "gry", "grn","hzl", "oth")) return(0)
    if (! grepl("^[0-9]{9}$", vals["pid"])) return(0)
    return(1)
  }
) %>% sum()
```
