---
title: "{hexmake} is one of the 5 Grand Prizes of the 2020 Shiny Contest"
post_date: 2020-07-15
layout: single
permalink: /hexmake-shiny-contest/
categories: r-blog-en
output: jekyllthat::jekylldown
excerpt_separator: <!--more-->
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.path="../assets/img/", eval = TRUE, error =TRUE, comment = "")
```

Hey y'all! 

I'm very happy to announce that my `{hexmake}` application has won one of the 5 Grand Prizes of the 2020 shiny contest, organized by RStudio!

`{hexmake}` is a pretty simple application when it comes to its idea: building hex stickers. 
But I wanted this simple idea to become a playground for showcasing more advanced `{shiny}` features: namely manipulating images with live display, importing and exporting data with a specific file format (here `.hex`), and importing and exporting to an external database (here a MongoDB). 

You can give `{hexmake}` a try at [connect.thinkr.fr/hexmake](https://connect.thinkr.fr/hexmake/), and read more about the Shiny Contest results on the [RStudio blog](https://blog.rstudio.com/2020/07/13/winners-of-the-2nd-shiny-contest/).
The source code of the application is at [github.com/ColinFay/hexmake](https://github.com/ColinFay/hexmake).

You can also play with `{hexmake}` on your machine by installing it with:

```{r eval = FALSE}
# Install
remotes::install_github("ColinFay/hexmake")
# Run
hexmake::run_app(with_mongo = FALSE)
```

If you want to use a mongodb as a back-end, you can launch one using docker with:

```{bash eval = FALSE}
docker run \
  -v /mongo/data/db:/data/db \
  -v /mongo/data/dump:/dump \
  -p 12334:27017 \
  -d --name mongohexmake \
  -e MONGO_INITDB_ROOT_USERNAME=myuser \
  -e MONGO_INITDB_ROOT_PASSWORD=mypassword \
  mongo:3.4 
```

Then launch the app using:

```{r eval = FALSE}
# Change these env variables to suite your mongo configuration
Sys.setenv("MONGOPORT" = 12334)
Sys.setenv("MONGOURL" = "127.0.0.1")
Sys.setenv("MONGODB" = "hex")
Sys.setenv("MONGOCOLLECTION" = "make")
Sys.setenv("MONGOUSER" = "myuser")
Sys.setenv("MONGOPASS" = "mypassword")
hexmake::run_app(with_mongo = TRUE)
```

