---
title: "Advent of Code 2021-02 with R & JavaScript"
post_date: 2021-12-02
layout: single
permalink: /aoc-2021-02/
categories: r-blog-en
output: jekyllthat::jekylldown
excerpt_separator: <!--more-->
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.path = "../assets/img/", eval = TRUE, error = TRUE, warning = FALSE)
```

Solving Advent of Code 2021-02 with R and JavaScript. 

Like yesterday, trying to write JS code that looks like R code.

## Disclaimer

[Disclaimer] Obviously, this post contains a big spoiler about Advent of Code, as it gives solutions for solving 2021-02.

[Disclaimer bis] I'm no JavaScript expert so this might not be the perfect solution. TBH, that's also the case for the R solution. 

## About the JavaScript code

The JavaScript code has been written in the same RMarkdown as the R code. It runs thanks to the `{bubble}` package: https://github.com/ColinFay/bubble.

```{r}
bubble::set_node_engine()
```

## Instructions

### One

+ We start at a position [0, 0] (horizontal, depth), then 
  - forward X increases the horizontal position by X units
  - down X increases the depth by X units
  - up X decreases the depth by X units

Result is horizontal * depth.

### Two

+ We start at a position [0, 0, 0] (horizontal, depth,), then 
  - down X increases your aim by X units
  - up X decreases your aim by X units
  - forward X does two things :
    - It increases your horizontal position by X units
    - It increases your depth by your aim multiplied by X

Result is horizontal * depth.

Find the complete instructions at: <https://adventofcode.com/2021/day/2>. 

## Reading inputs

```{r}
# Read
ipt <- read.delim("2021-02-aoc.txt", header = FALSE)$V1
```

```{node results = 'hide'}
const fs = require('fs')
var ipt = fs.readFileSync("2021-02-aoc.txt", 'utf8').split("\n").filter(x => x.length != 0);
```

## Part one 

### R solution

```{r}
# Horizontal and depth
start_pos <- c(horizontal = 0, depth = 0)
for (moves in ipt) {
  move <- strsplit(moves, " ")[[1]]

  if (move[1] == "up") {
    start_pos[2] <- start_pos[2] - as.numeric(move[2])
  }
  if (move[1] == "down") {
    start_pos[2] <- start_pos[2] + as.numeric(move[2])
  }
  if (move[1] == "forward") {
    start_pos[1] <- start_pos[1] + as.numeric(move[2])
  }
}

start_pos[1] * start_pos[2]
```

### JS solution

```{node results = 'hide'}
var start_pos = [0, 0]

// Gotcha: R for(x in y) is JS for(x of y)

for (moves of ipt) {
  // In JS, object have methods, so the split function
  // is containes inside the string object
  move = moves.split(' ')

  // Small gotcha here, JS i 0 based index
  if (move[0] == "up") {
    // R as.numeric is parseInt()
    start_pos[1] = start_pos[1] - parseInt(move[1])
  }
  if (move[0] == "down") {
    start_pos[1] = start_pos[1] + parseInt(move[1])
  }
  if (move[0] == "forward") {
    start_pos[0] = start_pos[0] + parseInt(move[1])
  }
}
```

```{node}
start_pos[0] * start_pos[1]
```

## Part two 

### R solution

```{r}
# Horizontal, depth, and aim
start_pos <- c(horizontal = 0, depth = 0, aim = 0)
for (moves in ipt) {
  move <- strsplit(moves, " ")[[1]]

  if (move[1] == "up") {
    start_pos[3] <- start_pos[3] - as.numeric(move[2])
  }
  if (move[1] == "down") {
    start_pos[3] <- start_pos[3] + as.numeric(move[2])
  }
  if (move[1] == "forward") {
    start_pos[1] <- start_pos[1] + as.numeric(move[2])
    start_pos[2] <- start_pos[2] + (start_pos[3] * as.numeric(move[2]))
  }
}

start_pos[1] * start_pos[2]
```

### JS solution

```{node results = 'hide'}
var start_pos = [0, 0, 0]

for (moves of ipt) {
  move = moves.split(' ')
  if (move[0] == "up") {
    start_pos[2] = start_pos[2] - parseInt(move[1])
  }
  if (move[0] == "down") {
    start_pos[2] = start_pos[2] + parseInt(move[1])
  }
  if (move[0] == "forward") {
    start_pos[0] = start_pos[0] + parseInt(move[1])
    start_pos[1] = start_pos[1] + (start_pos[2] * parseInt(move[1]))
  }
}
```

```{node}
start_pos[0] * start_pos[1]
```