---
title: "Advent of Code 2021-01 with R & JavaScript"
post_date: 2021-12-01
layout: single
permalink: /aoc-2021-01/
categories: r-blog-en
output: jekyllthat::jekylldown
excerpt_separator: <!--more-->
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.path = "../assets/img/", eval = TRUE, error = TRUE, warning = FALSE)
bubble::set_node_engine()
```

Solving Advent of Code 2021-01 with R and JavaScript. 

I'll try in this post to write R code that looks like the JavaScript one, using `{purrr}`, to show how both syntaxes can look alike.
The idea is to highlight how you can leverage your `{purrr}` knowledge when writting JavaScript code.

## Disclaimer

[Disclaimer] Obviously, this post contains a big spoiler about Advent of Code, as it gives solutions for solving day 1.

[Disclaimer bis] I'm no JavaScript expert so this might not be the perfect solution. TBH, that's also the case for the R solution. 

## About the JavaScript code

The JavaScript code has been written in the same RMarkdown as the R code. It runs thanks to the `{bubble}` package: https://github.com/ColinFay/bubble

## Instructions

+ Count the number of times n is larger that n-1

+ Count the number of times n + n-1 + n-2 is larger that n-1 + n-2 + n-3

Find the complete instructions at: <https://adventofcode.com/2021/day/1>. 

## Reading inputs

```{r}
# Read
library(purrr)
ipt <- read.delim("2021-01-aoc.txt", header = FALSE)$V1
```

> // Reading the file and converting it to int

```{node results = 'hide'}
const fs = require('fs')
var ipt = fs.readFileSync("2021-01-aoc.txt", 'utf8').split("\n").filter(x => x.length != 0);
ipt = ipt.map(x => parseInt(x));
```

## Part one 

### R solution

```{r}
2:length(ipt) |>
  map_dbl(
    ~ ipt[.x] - ipt[.x - 1]
  ) |>
  keep(
    ~ .x > 0
  ) |>
  length()
```

### JS solution

```{node}
ipt.map( 
  // inside map() in JS, the second argument to the anonymous
  // function is the index inside the array. 
  // It mimics what we have in map(2:length(ipt)) in R 
  (x, y) => ipt[y] - ipt[y-1] 
  ).filter(
    // filter() will keep the element that matches the anonymous
    // predicate function
    x => x >= 0
    ).length;
```


### Part two 

### R solution

```{r}
tmp <- 3:length(ipt) |>
  map_dbl(
    ~ ipt[.x] + ipt[.x - 1] + ipt[.x - 2]
  )

2:length(tmp) |>
  map_dbl(
    ~ tmp[.x] - tmp[.x - 1]
  ) |>
  keep(
    ~ .x > 0
  ) |>
  length()
```

### JS solution

```{node}
var tmp = ipt.
  map( 
    // Creating n + n-1 + n-2
    (x, y) => ipt[y] + ipt[y-1] + ipt[y-2]
    );

tmp.
  map( 
    // Same as part 1
    (x, y) => tmp[y] - tmp[y-1] 
    ).
  filter(
    // keeping the one that have raised
    x => x > 0
  ).
  length;
```
